---
title: "The basics, with two models"
author: "Ricardo Mayer"
date: "12/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r source-and-lib, message=FALSE, warning=FALSE}
source('./R/combinations_functions.R')
```


## Estimating one VAR model to predict GDP growth

We will start with a ready-to-use data set of quarterly data ready to use in our VAR estimation and forecast. In particular, we will *not* discuss the following important points:
    - how the data was obtained
    - how monthly data was converted to quaterly frequency
    - how monthly data was *extended* to complete its current final quarter
    - how (potentially) exogenous data was forecasted in order to make it available to produce conditional forecasts
    
All those points are discussed in the data preparation document, see *here*

    
    


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
