---
title: 'Working with multiple VARs'
author: "Ricardo Mayer"
date: "12/6/2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r source-and-lib, message=FALSE, warning=FALSE}
source('./R/combinations_functions.R')
```


## VAR-ready data set
    
We will use the example dataset with domestic series from Uruguay and few external series. The rds file contains the country's name, the transformation applied to rgdp to render it stationary and two data sets: the original or raw data and one ready to used in VAR estimation,  containing only stationary versions of the original series. 


```{r loading_data}
data_object_ury <- readRDS("./data/examples/example_data_ury.rds")
country <- data_object_ury$country_name
target_transformation <- data_object_ury$target_transformation
raw_data <- data_object_ury$raw_data
var_data <- data_object_ury$transformed_data
print(paste0("target transformation: ",target_transformation))

names_exogenous <- c("ip_us","ip_asia","ip_ue","ip_bra","act_eco_bra","emae_arg")

extension_of_exo_list <- readRDS(file = "./data/examples/example_extension_of_exo.rds")
extension_of_exo <- extension_of_exo_list$extended_exo

cv_extension_of_exo_list <- readRDS(file = "./data/examples/example_cv_extension_of_exo.rds")
cv_extension_of_exo <- cv_extension_of_exo_list$future_exo_cv

```



## Getting a tibble with multiple specifications

Throught this document we will work with just a few models of size 3 (i.e. the target variable plus two more variables), to keep things simple and relatively fast. But there nonetheless enough models to illustrate all the important points, including the reduction of the set of considered models.

The strategy is to generate all possible specifications given our choices and the select randomly 20 combinations of variables and lags. We show you how to select 20 models at random but for reproducibility we will load a previosly saved selection.

In particular, the all specifications are generated:
- using all variables in the data set
- forming all size-3 combinations of variables (target variable and two more)
- using two choices of maximum lag, 3 and 5
- using the VARselect function from *vars* we add any other choice of max lag suggested by various information criteria
- This generates 1576 variables-and-lag combinations 
- For each variables-and-lag combination we consider three version of the model: the unrestricted one and two restricted versions, one using a threshold of 1.65 and other using a threshold of 2.0
- Finally we select, randomly,  20 variables-and-lag combinations (and their three restricted/unrestricted versions)

```{r subsetsofspecifications}
var_size <- 3 
all_variables <- colnames(var_data)
target_variable <- "rgdp"
non_target_fixed <- c("")

# only "size 3" specifications
specifications_all <- all_specifications(
  var_size = 3, all_variables = colnames(var_data),
  lag_choices = c(3, 5), use_info_lags = TRUE,
  var_data = var_data, t_thresholds = c(1.65, 2), names_exogenous = names_exogenous)

# specifications_n20 <- dplyr::sample_n(specifications_all, 20)
# saveRDS(object = specifications_n20, file = "./data/examples/ury_specifications_n20.rds")

specifications_n20 <- readRDS(file = "./data/examples/ury_specifications_n20.rds")

specifications_n20endo <- specifications_n20[-c(6, 9, 13, 18, 19), ]

```

Looking at a table with values of max lag in our specifications, we found that unsurprignly 3 and 5 are the most common values with 435 appeareances each, that max lag 1 resulted very popular after applying VARselect and max lag 4 was also pretty popular. Values 2, 6 and 7 maxlag do appear but much fewer cases.

Our selection of 20 rows, it happens to preserve that structure by picking only maxlag values of 1, 3, 4 and 5.

```{r large_and_small_tbl_of_specs}
nrow(specifications_all)
table(specifications_all$lags)
table(specifications_n20$lags)


```

A closer look to the variables in these specifications, shows us that both edogenous and exogenous variables appear in some of the VARs

```{r see_spec_20}
knitr::kable(specifications_n20)
```



## Fit all specifications and keep only acceptable ones

```{r  passing_models_size3small, cache=TRUE}
ftmt_n20endo <- fit_tests_models_table(specifications_n20endo, var_data = var_data, names_exogenous = names_exogenous)
ftmt_n20oneexo <- fit_tests_models_table(specifications_n20oneexo, var_data = var_data, names_exogenous = names_exogenous)
ftmt_n20twoexo <- fit_tests_models_table(specifications_n20twoexo, var_data = var_data, names_exogenous = names_exogenous)
ftmt_n20threeexo <- fit_tests_models_table(specifications_n20threeexo, var_data = var_data, names_exogenous = names_exogenous)
ftmt_n20 <- fit_tests_models_table(specifications_n20, var_data = var_data, names_exogenous = names_exogenous)

# foo1 <- ftmt_n20oneexo$passing_models
# foo2 <- ftmt_n20oneexo$tried_models
# foo3 <- ftmt_n20oneexo$n_lost_to_threshold
# foo4 <- ftmt_n20oneexo$n_lost_to_roots
# foo5 <- ftmt_n20oneexo$n_lost_to_white

foo1 <- ftmt_n20oneexo$passing_models
foo2 <- ftmt_n20oneexo$tried_models
foo3 <- ftmt_n20oneexo$n_lost_to_threshold
foo4 <- ftmt_n20oneexo$n_lost_to_roots
foo5 <- ftmt_n20oneexo$n_lost_to_white

pm_n20 <- ftmt_n20[["passing_models"]]
print(head(pm_n20))


totasia165 <- pm_n20[15, ]


```


## Add forecast performance information via Time Series Cross Validation 


```{r tscv, message=FALSE, warning=FALSE, results='hide', cache=TRUE}
n_cv <- 10
fc_horizon <- 8
training_length <- "per_cv_maxs" 
target_transform <- target_transformation
target_level_ts <- na.omit(raw_data[, target_variable])


cv_n20 <- cv_var_from_model_tbl(h = fc_horizon,
                             training_length = training_length, 
                             n_cv = n_cv,
                             models_tbl = pm_n20, 
                             var_data = var_data, 
                             fit_column = NULL, 
                             target_transform = target_transform,
                             target_level_ts = target_level_ts, 
                             names_exogenous = names_exogenous, 
                             future_exo = extension_of_exo, 
                             extended_exo_mts = cv_extension_of_exo,
                             keep_varest_obj = FALSE
                             )

pm_n20endo <- ftmt_n20endo[["passing_models"]]
cv_n20endo <- cv_var_from_model_tbl(h = fc_horizon,
                             training_length = training_length, 
                             n_cv = n_cv,
                             models_tbl = pm_n20endo, 
                             var_data = var_data, 
                             fit_column = NULL, 
                             target_transform = target_transform,
                             target_level_ts = target_level_ts, 
                             names_exogenous = names_exogenous, 
                             future_exo = extension_of_exo, 
                             extended_exo_mts = cv_extension_of_exo,
                             keep_varest_obj = FALSE
                             )


# 
# tic()
# cv_n20_usefit <- cv_var_from_model_tbl(h = fc_horizon,
#                                                training_length = training_length, 
#                                                n_cv = n_cv,
#                                                models_tbl = pm_n20, 
#                                                var_data = var_data, 
#                                                fit_column = "fit", 
#                                                target_transform = target_transform,
#                                                target_level_ts = target_level_ts, 
#                                                names_exogenous = names_exogenous, 
#                                                future_exo = extension_of_exo, 
#                                                extended_exo_mts = cv_extension_of_exo,
#                                                keep_varest_obj = FALSE
# )
# toc()


```

## Combine individual-models forecasts into a weigthed average forecast

### First step: forecasts from individual models

```{r full_sample_fc, results='hide', cache=TRUE}


fcs_from_endo <- forecast_var_from_model_tbl(
  models_tbl = cv_n20endo,
  var_data = var_data, 
  fc_horizon = fc_horizon, 
  target_transform = target_transform,
  target_level_ts = target_level_ts,
  names_exogenous = names_exogenous, 
  extended_exo_mts = extension_of_exo, 
  keep_wide_tbl = TRUE
)


fcs_from_all <- forecast_var_from_model_tbl(
  models_tbl = cv_n20,
  var_data = var_data, 
  fc_horizon = fc_horizon, 
  target_transform = target_transform,
  target_level_ts = target_level_ts,
  names_exogenous = names_exogenous, 
  extended_exo_mts = extension_of_exo, 
  keep_wide_tbl = TRUE
)

long_tbl <- fcs_from_all$models_tbl
wide_tbl <-  fcs_from_all$models_tbl_wide
info_per_h <- fcs_from_all$models_info_per_h

```

### Second step: do a weighted average of forecasts' means

```{r ensemble_mean}
# h_max <- fc_horizon
ensemble_fc_list <- fc_mean_of_VAR_ensemble(models_tbl = fcs_from_all$models_tbl)
ensemble_fc_tbl <- ensemble_fc_list$ensemble_tbl
ensemble_fc_ts <- ensemble_fc_list$weighted_avg_fc_yoy

```

### Third step: TSCV of weighted average, compute RMSE and put all together

```{r tscv_of_ensemble, results='hide', cache=TRUE}
ensemble_cv <- cv_of_VAR_ensemble(var_data = var_data,
                                  used_cv_models = fcs_from_all$models_tbl_wide,
                                  fc_horizon = fc_horizon,
                                  n_cv = n_cv,
                                  training_length = training_length,
                                  cv_extension_of_exo = cv_extension_of_exo,
                                  names_exogenous = names_exogenous,
                                  max_rank_h = NULL,
                                  target_transform = target_transform,
                                  target_level_ts = target_level_ts)

```

and bind them

```{r bind_indiv_ensemble}

ensemble_fc_and_rmse <- ensemble_fc_tbl %>% 
  mutate(rmse_h = paste0("rmse_", 1:n()),
         rmse = ensemble_cv$ensemble_rmse,
         rank = -1)

fcs_models_to_bind <- fcs_from_all$models_tbl %>% 
  mutate(lags = list(lags)) %>% 
  dplyr::select(names(ensemble_fc_and_rmse))


models_and_ensemble_fcs <- rbind(ensemble_fc_and_rmse, 
                                 fcs_models_to_bind)

```


### Three steps in one function

```{r threestepsinone, results='hide', cache=TRUE}

final_tbl <- ensemble_fc_from_models_rmse(models_tbl_with_rmse = cv_n20,
                             var_data = var_data, 
                             n_cv = n_cv, 
                             training_length = training_length, 
                             max_rank_h = NULL, 
                             fc_horizon = fc_horizon,
                             names_exogenous = names_exogenous, 
                             target_transform = target_transform,
                             target_level_ts = target_level_ts,
                             extension_of_exo = extension_of_exo, 
                             cv_extension_of_exo = cv_extension_of_exo, 
                             fit_column = NULL)
```



## Change the number of models based on model rankings


```{r fcs_max_10, results='hide'}

foo_long <- discard_by_rank(fcs_from_all$models_tbl, max_rank_h = 3, is_wide = FALSE)

foo_wide <- discard_by_rank(fcs_from_all$models_tbl_wide, max_rank_h = 3, is_wide = TRUE)

sort(unique(foo_long$short_name))
sort(unique(foo_wide$short_name))
identical(sort(unique(foo_long$short_name)), sort(unique(foo_wide$short_name)))



fcs_from_10 <- forecast_var_from_model_tbl(
  models_tbl = cv_n20,
  var_data, 
  fc_horizon, 
  target_transform = target_transform,
  target_level_ts = target_level_ts,
  names_exogenous = names_exogenous, 
  extended_exo_mts = extension_of_exo, 
  keep_wide_tbl = TRUE,
  max_rank_h = 10
)



final_tbl_10 <- ensemble_fc_from_models_rmse(
                             models_tbl_with_rmse = cv_n20,
                             var_data = var_data, 
                             n_cv = n_cv, 
                             training_length = training_length, 
                             max_rank_h = 10, 
                             fc_horizon = fc_horizon,
                             names_exogenous = names_exogenous, 
                             target_transform = target_transform,
                             target_level_ts = target_level_ts,
                             extension_of_exo = extension_of_exo, 
                             cv_extension_of_exo = cv_extension_of_exo, 
                             fit_column = NULL)


```






The function forecast_var_from_model_tbl by default, it just take all specifications in given tibble and produce forecasts of the target variable, more precisely it creates a list-column containing mean forecasts of YoY growth of the target variable over the specified forecast horizon. Some options include to keep varest or forecast objects (not advised with very long tibbles) and more importantly it can restrict the sample of models by keeping only the best  models at each horizon and computing a (rmse-)weighted average mean forecast.

In this case we take our models from the cv-exercises, thus all models where already tested for stability, restrictions and white noise residuals and therefore we can spare those test here, which is the default behaviour.


this_fc_tbl

